### 2009-12-19 trunk/Papers/reasoning-challenges/conclusion.tex
changes to abstract, intro, conclusions

>>>>

%\todo{this is cut and paste from Guido's text, which I like but which
%  was in a different place. The conclusion should be cleaned up and
%  made real text. - ak}

The new direction we propose for the IPC has significant advantages for
each of the competition's three main groups of stakeholders: the
organisers, the participants, and the audience.  Organisers benefit from a
mostly automated platform that is maintained by the community, which should
greatly reduce the effort needed to run a competition. Participants benefit
from a well-defined environment for running their systems, short feedback
times for system evaluation, and a standardised test and benchmark platform
for experimenting with novel algorithms. The audience -- in this case the
wider planning and scientific communities -- benefits from improved
discussion and reporting facilities. Since the scientific value of
competitions is to find out \emph{why} a system performs the way it does,
not only \emph{that} it performs a certain way, enabling users to
contribute to the competition process reduces the chance that essential
avenues of investigation and evaluation are missed. Instead, a robust and
active competition community only serves to enhance the wider planning
community as a whole.

%This new approach has advantages for each of the three groups of
%people involved in a competition, the organisers, the participants,
%and the audience (which usually includes organisers and
%participants). \todo{do we want to use the term ``stakeholders'' for
%  all of these people together? - ak} %RP: I like stakeholders.

%\todo{I don't think we should have acknowledgments in the anonymous
%  version. - ak}
%RP: I agree; the acknowledgements should be left out to anonymise the
%paper

%\subsubsection*{Acknowledgements}
%The original idea to organize a ``rolling'' competition in the CP
%community is due to Christian Schulte and Peter Stuckey.

### 2009-12-19 trunk/Papers/reasoning-challenges/introduction.tex
changes to abstract, intro, conclusions

This paper proposes a new direction for the IPC, building on the strength
of previous competitions, but centred around a novel online infrastructure.
Most notably, we envision a more interactive and potentially more
wide-ranging competition featuring:
%
\begin{itemize}
\item A ``rolling'' competition format with community-initiated track and
problem specifications,

@@

\item A centralised repository offering researchers access to the latest
domains and planners, and a forum for community news, discussion, and
interaction.
\end{itemize}

@@

From a technical point of view, it is important to note that such an idea
would not have been practical ten years ago during the time of the first
IPC. It is only recently, due to advances in computing hardware and network
bandwidth, that a scenario such as this is now feasible. Even so, our
proposed infrastructure nevertheless presents some technical challenges
that must be addressed (e.g., security, , accessibility, maintenance,
etc.). However, we believe that for the most part, solutions to these
problems can already be found by looking at similar (although smaller
scale) competition efforts in the wider research community.

Most importantly, we believe this proposal offers significant benefits to
competition organisers, participants, and the wider planning community.
\textbf{[RP: say something more here]} Furthermore, since much of the
infrastructure we describe here is not tied to a specific research area,
attempts to establish similar competitions could benefit from this
approach, leading to joint ventures with other research fields---and the
possibility of future inter-disciplinary challenges.

In the remainder of the paper we outline our proposal for a new competition
infrastructure, and the benefits we believe this approach could bring to
the planning community and beyond.

>>>>

This paper proposes a new direction for the IPC, building on the strength
of previous competitions and inspired by competitions in other research
communities. Most notably, we envision an interactive and user-centric
%potentially more wide-ranging
competition, focused around a novel online infrastructure featuring:
%Most notably, we envision a more interactive and potentially more
%wide-ranging competition featuring:
%
\begin{itemize}
\item A ``rolling'' competition format with community-initiated track and
problem specifications,

@@

\item A centralised repository offering researchers access to the latest
domains and planners, and a social networking forum for community news,
discussion, and interaction.
\end{itemize}

@@

From a technical point of view, it is important to note that such an idea
would not have been practical ten years ago during the time of the first
IPC. It is only recently, due to advances in computing hardware and network
bandwidth, that a scenario such as this is now feasible. Even so, our
proposed infrastructure nevertheless presents some technical challenges
that must be addressed (e.g., security, accessibility, maintenance,
etc.). However, we believe that for the most part, solutions to these
problems can already be found by looking at similar (although smaller
scale) competition efforts in the wider research community.

Most importantly, we believe this proposal offers significant benefits to
the competition stakeholders: organisers, participants, and the wider
planning community. Furthermore, since much of the infrastructure we
describe is not tied to a specific research area, attempts to establish
similar competitions could benefit from this approach, leading to joint
ventures with research fields beyond planning---and the
possibility of future inter-disciplinary challenges.

%RP: Alexander's first couple sentences in the next section are a good lead
%in to the content of the paper. I don't think we need much or anything
%as a transition here, especially if we're short on space.

%In the remainder of the paper we outline our proposal for a new competition
%infrastructure, and the benefits we believe this approach could bring to
%the planning community and beyond.

### 2009-12-19 trunk/Papers/reasoning-challenges/levels.tex
changes to abstract, intro, conclusions

We present our proposal for a new planning competition infrastructure
in terms of an ascending scale of levels, each of which makes it
easier or more rewarding for the community to participate in the
competition process.  As we will discuss below, we believe that the
competition format we propose is applicable in fields beyond planning
as well.  Therefore we will talk about \emph{systems}, \emph{problem
  classes}, and \emph{problem instances} instead of planners, planning
domains, and problems.

@@

As background and starting point of our proposal, consider the process
that underlies the current IPC competition.  Every two years, the IPC
committee appoints one or two organizers for each track of that year's
IPC competition.  These organizers assemble a number of problem
classes and instances, by developing them themselves or soliciting
them from the IPC committee or the community at large; sometimes this
requires extending the Planning Domain Description Language (PDDL,
\cite{PDDL}) to more advanced classes of problems.  After the
participants have developed their planners, they submit the source
code to the organizers, who then compile and run the planners on the
test machines.  Finally, the organizers prepare and present a report,
and the best-performing systems win awards.

\todo{Ron, can you fix this so it becomes true? - ak}

\subsection{Level 1: Collecting problems}

While this format for the IPC has been highly successful in the past
ten years, there is also a number of ways in which it can be improved
and streamlined.  One such direction concerns the way in which the
problem classes on which an IPC runs are collected.  There is no doubt
that the selection of IPC domains is a fundamental driving force for
the development of new planning algorithms.  Unfortunately, the
current process makes it difficult for users of planning technology,
who may come from a different community, to get a new
application-motivated domain accepted as an IPC benchmark.  This is
unsurprising, given that an IPC organizer must be able to justify
their selection of domains, and picking untested domains from other
communities is not the best way to do this.  However, this leads to a
selection process that is less dynamic and open to suggestions from
users than it could be.  A converse movement is that to keep the
number of domains that must be tested in each round of the competition
manageable, older domains are sometimes ``retired'', i.e.\ removed
from subsequent competitions.  This comes with a risk of a false sense
of progress, in which retired domains are remembered as ``solved''
although they were removed for completely independent reasons.
Finally, it has been argued \cite{ToughNuts} that some IPC organizers
tend to favor the development of new types of planning problems and
new dialects of PDDL over trying to find challenging benchmark
problems for existing versions of PDDL.

We propose that one way in which these problems can be alleviated is a
web repository in which any user can upload problem classes.  This
takes up Hoffmann's \shortcite{ToughNuts} idea to create a
community-driven repository for ``tough nuts'' -- i.e., instances that
were shown to be hard to solve in the past --, and takes it one step
further: In our proposal, \emph{all} IPC problem classes are suggested
by the community, much in line with the proposal by Roberts and Howe
\shortcite{roberts07:_call_for_partic}. This approach offers a
principled, publicly documented and accessible avenue for proposing
new problem classes, which benefits both application-oriented users of
planning technology and planning researchers who come up with tough
challenges for existing PDDL dialects. It also makes it easy to record
retired problem classes alongside more current ones, and makes it easy
to keep them accessible.

Submitting a problem class should involve three pieces of data. First,
the user must submit a collection of instances of the problem class or
a generator for such instances.  Second, the user must specify some
metadata including the class name and a human-readable description,
maybe with links to the application literature, complexity results,
and so on.  Third, the user must specify the \emph{type} of the
problem class, which determines what systems the data is suitable for
-- i.e., the language in which the problem is specified (such as a
particular dialect of PDDL or EUROPA \todo{cite}) and the type of
planning problem this is (propositional, temporal, uncertainty,
etc.).  If it is possible to convert between one specification
language and another automatically, this conversion could be performed
each time a problem is submitted.

Problem classes and instances can be submitted by any registered
user. In order to avoid abuse, the number of instances should be
limited (e.g.\ per hour), and a simple validity check should be
possible (such as running a parser if the format is PDDL).

@@

But given that all the domains are available on a server, why stop
there?  We propose to make it possible to upload the \emph{planners}
to the same server as well, and then continuously and automatically
run the planners on the domains.  While the ability to upload domains
in Level 1 represents a conservative extension of the existing IPC
infrastructure, it is here that we depart more radically from the
existing format: Instead of a regular competition with organizers who
run the planner evaluation every other year, we propose a rolling
evaluation scheme, in which the current evaluation results would be
available online at any given time.

@@

We propose that every newly submitted planner is run automatically on
all available problem instances, and every newly submitted problem
instance is sent to all available planners.  The scheduling of this
process can be under the control of an automated system for continuous
integration, such as Hudson.\footnote{\url{http://hudson-ci.org/}}
Each run of a planner on a problem instance or problem class can take
place in a fresh instance of the virtual machine image.  The
evaluation infrastructure then inserts the problem data and the
planner into the virtual machine, runs the planner, measures
performance statistics (including which problems they could solve and
how much time this took them; see below), and stores the results in a
database.  Users will be able to view the results in dynamically
generated reports on the main website; there may also be functionality
to customize these reports to focus on certain types of problems.
Because all planners are run on all problem instances, problems in
which only a small and somewhat arbitrary selection of instances of a
certain problem are actually used in the evaluation \cite{pg2008002}
can be avoided.

In order to minimize security risks -- which arise when user-submitted
programs are being run automatically --, the individual virtual
machines need to be sandboxed quite tightly.  There is no need for an
individual virtual machine to be connected to any network, beyond
being able to enter performance measurements into a database.  By
further imposing resource limits on the number of available
processors, memory and processor time, and disk space, we believe the
potential for abuse is limited.  It can be limited further by only
allowing planner uploads to registered users.

While we see our proposed system's ability to generate up-to-date
reports automatically as a major strength, there may still be a
benefit in maintaining the current style of regular competitions at
discrete timepoints, e.g.\ to encourage the development of faster
planners through awards.  Such a competition could easily be based on
our system, by announcing a deadline and a specified subset of problem
classes. The selection of this subset could be informed, among other
things, by the report on problems that many planners can't
solve. Participants could then upload their planners at their leisure;
``running'' the competition amounts to taking a snapshot of the
performance reports on the selected problems when the deadline comes
and announcing the results.  By recording timestamps whenever a
performance measurements is made, this snapshot could remain available
on the website simply as one special report for a certain point in
time.  In such a case, the authors of planners should have the option
of marking them as ``non-competing'' to exclude them from the
competition snapshot, and maybe also from the up-to-date performance
reports.

@@

The obvious criticism that can be brought up against this proposal is
that of cost -- both in terms of implementing the necessary software
and in terms of running the hardware on which the continuous
evaluation is performed.

@@

One advantage of the evaluation infrastructure proposed above is that
its applicability is not limited to the evaluation of planners.  Many
reasoning tasks, including constraint solving and theorem proving, are
being evaluated by running solvers over problem instances and
reporting performance measures.  At the level of abstraction at which
our continuous evaluation system could be implemented, there is no
fundamental difference between the problem type ``propositional
planning problem in PDDL 1.0'' and the problem type ``\todo{something
  from CP or similar}'' -- in both cases, the system can collect
problem classes and instances of this type and run them through
solvers that declare themselves to accept such inputs.  This means
that there is a potential for sharing development and maintenance
costs for the evaluation platform among different communities.

However, there are of course limitations to the type of problem for
which it makes sense to run a community-driven continuous evaluation.
For one thing, it must be acceptable in the community that every
system runs on the same hardware and software configuration, as
implemented in the virtual machine images.  On the other hand, our
system makes most sense where the ratio of problem classes to systems
is relatively high.  Neither of the two is true, for instance, for
most natural language processing competitions: Different research
groups tend to use a variety of programming languages and development
environments to build their systems, and there are typically many more
systems in any competition than there are problem classes (corpora).
However, our approach is completely consistent with the requirements
of a community where solving a single hard problem instance for the
first time counts as an achievement, such as in theorem proving.

>>>>

We present our proposal for a new planning competition infrastructure
in terms of an ascending scale of levels, each of which makes it
easier or more rewarding for the community to participate in the
competition process. As we will discuss below, we believe that the
competition format we propose is applicable to fields beyond planning.  We
will therefore refer to \emph{systems}, \emph{problem classes}, and
\emph{problem instances}, instead of planners, planning domains, and
planning problems.

@@

As background, and a starting point for our proposal, consider the process
that underlies the current IPC. Every two years, the IPC committee looks
for volunteers to organize and run the various tracks of that year's
competition. These organizers are responsible for assembling a number of
problem classes and instances, by developing them themselves or soliciting
them from the wider community. Sometimes this process results in, or
coincides with, a new dialect of the Planning Domain Description Language
(PDDL, \cite{PDDL}), which describes for participants the supported
language for that year's competition. After the participants have developed
their systems, they submit the source code to the organizers, who then
compile and run the planners on the test machines.  Performance data is
collected and analysed.  Finally, the organizers prepare and present a
report, and the best-performing systems win awards.

\subsection{Level 1: Problem collection}

While this format for the IPC has been highly successful in the past
ten years, there is also a number of ways in which it can be improved and
streamlined.  One such direction concerns the way in which problem classes
are collected.  There is no doubt that the existing collection of IPC
domains is a fundamental driving force for the development of new planning
algorithms.  Unfortunately, the current selection process makes it
difficult for users of these systems, who may come from different
communities, to get a new application-motivated domain accepted as an IPC
benchmark.  This is unsurprising, given that IPC organizers must be able to
justify their selection of domains, and picking untested domains from other
communities is not the best way to do this.  However, this leads to a
selection process that is less dynamic than it could be, and more
restrictive for users wanting to contribute suggestions.  Conversely, to
keep the number of domains that must be tested in each round of the
competition manageable, older domains are sometimes ``retired'', i.e.\
removed from subsequent competitions.  This comes with a risk of a false
sense of progress, in which retired domains are remembered as ``solved''
although they were removed for completely independent reasons.  Finally, it
has been argued \cite{ToughNuts} that some IPC organizers tend to favor the
development of new types of planning problems and new dialects of PDDL over
trying to find challenging benchmark problems for existing versions of
PDDL. Such approaches run the risk of ``reinventing'' past problem classes
(albeit in different guises), or redirecting investigation away from
current challenges.

One way to alleviate these problems is by introducing a web repository
where users can upload problem classes. This revisits Hoffmann's
\shortcite{ToughNuts} idea to create a community-driven repository for
``tough nuts'' -- i.e., instances that were shown to be hard to solve in
the past -- and takes it one step further: in our proposal, \emph{all} IPC
problem classes are provided by the community, much in line with the
proposal by Roberts and Howe \shortcite{roberts07:_call_for_partic}. Such a
repository offers a principled, publicly documented and accessible avenue
for proposing new problem classes, which benefits both application-oriented
users of planning technology and researchers who invent tough challenges
for existing PDDL dialects. It also makes it easy to record retired problem
classes alongside active ones, and keep them all accessible.

Problem classes and instances can be submitted by any registered user. 
Submitting a problem class should involve three pieces of data. First,
the user should submit a collection of instances of the problem class or
a generator for such instances.  Second, the user should specify some
metadata including the class name and a human-readable description,
possibly with links to the applicable literature, complexity results, and
so on.  Third, the user should specify the \emph{type} of problem class,
which determines what systems the data is suitable for, i.e., the language
in which the problem is specified (such as a particular dialect of PDDL or
EUROPA \todo{cite}) and the type of planning problem (e.g., propositional,
temporal, uncertainty, etc.).  If it is possible to convert between one
specification language and another automatically, this conversion could be
performed each time a problem is submitted.
In order to avoid abuse, the number of instances should be limited (e.g.,
per hour), and a simple validity check should be possible (e.g., running a
parser to verify that the format is PDDL).

@@

Given that all the domains are available on a server, why stop
there?  We propose to make it possible to upload the \emph{planners}
to the same server as well, and then continuously and automatically
run the planners on the domains.  While the ability to upload domains
in Level 1 represents a conservative extension of the existing IPC
infrastructure, it is here that we depart more radically from the
existing format: instead of a regular competition with organizers who
run the planner evaluation every other year, we propose a rolling
evaluation scheme, in which the current evaluation results would be
available online at any given time.

@@

We propose that every newly submitted planner is run automatically on
all available problem instances, and every newly submitted problem
instance is sent to all available planners.  The scheduling of this
process can be under the control of an automated system for continuous
integration, such as Hudson.\footnote{\url{http://hudson-ci.org/}}
Each run of a planner on a problem instance or problem class can take
place in a fresh instance of the virtual machine image.  The
evaluation infrastructure then inserts the problem data and the
planner into the virtual machine, runs the planner, measures
performance statistics (including which problems they could solve and
how much time this took them; see below), and stores the results in a
database.  Users will be able to view the results in dynamically
generated reports on the main website; there may also be functionality to
customize these reports to focus on certain types of problems.  Because all
planners are run on all problem instances, and users can ensure that
challenging problem instances are included in the competition, the chance
of systems being evaluated on only a small and somewhat arbitrary selection
of certain problems \cite{pg2008002} can be avoided.

In order to minimize security risks, which can arise when user-submitted
programs are being run automatically, the individual virtual
machines need to be sandboxed quite tightly.  There is no need for an
individual virtual machine to be connected to any network, beyond
being able to enter performance measurements into a database.  By
further imposing resource limits on the number of available
processors, memory and processor time, and disk space, we believe the
potential for abuse is limited.  It can be limited further by only
allowing planner uploads to registered users.

While we see our proposed system's ability to generate up-to-date
reports automatically as a major strength, there may still be a
benefit in maintaining the current style of regular competitions at
discrete timepoints, e.g., to encourage the development of faster
planners through awards.  Such a competition could easily be based on
our system, by announcing a deadline and a specified subset of problem
classes. The selection of this subset could be informed, among other
things, by the report on problems that many planners can't
solve. Participants could then upload their planners at their leisure;
``running'' the competition amounts to taking a snapshot of the
performance reports on the selected problems when the deadline comes
and announcing the results.  By recording timestamps whenever a
performance measurements is made, this snapshot could remain available
on the website simply as one special report for a certain point in
time.  In such a case, the authors of planners should have the option
of marking them as ``non-competing'' to exclude them from the
competition snapshot, and maybe also from the up-to-date performance
reports.

@@

The obvious criticism that can be brought against this proposal is
that of cost -- both in terms of implementing the necessary software
and in terms of running the hardware on which the continuous
evaluation is performed.

@@

One advantage of the evaluation infrastructure proposed above is that
its applicability is not limited to the evaluation of planners.
%RP: I don't see why this needs to be restricted to "reasoning" tasks
Many reasoning tasks, including constraint solving and theorem proving, are
being evaluated by running solvers over problem instances and reporting
performance measures.  At the level of abstraction at which our continuous
evaluation system could be implemented, there is no fundamental difference
between the problem type ``deterministic planning problem in PDDL 1.0'' and
the problem type ``\todo{something from CP or similar}'' -- in both cases,
the system can collect problem classes and instances of this type and run
them through solvers that declare themselves to accept such inputs.  This
means that there is a potential for sharing development and maintenance
costs for the evaluation platform among different communities.

However, there are of course limitations to the type of problem for
which it makes sense to run a community-driven continuous evaluation.
For one thing, it must be acceptable in the community that every
system runs on the same hardware and software configuration, as
implemented in the virtual machine images.  On the other hand, our
system makes most sense where the ratio of problem classes to systems
is relatively high.  Neither of the two is true, for instance, for
most natural language processing competitions: different research
groups tend to use a variety of programming languages and development
environments to build their systems, and there are typically many more
systems in any competition than there are problem classes (corpora).
However, our approach is completely consistent with the requirements
of a community where solving a single hard problem instance for the
first time counts as an achievement, such as in theorem proving.
